---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# Register studies

<!-- badges: start -->
<!-- badges: end -->

Data utilised in register studies, often called as register-based data or administrative data, typically consists of id's, time stamp variables and codes describing different events. This package has been designed specifically for handling of such data. 

Package is based on tidyverse and examples utilise widely tidyverse programming style. The functions are designed with ICD codes in mind, but can be used with any kind of string formatted codes and are thus not necessarily limited to healthcare data. Users are allowed to make their own definitions of code classifiers, so the package suits needs of different countries having their own code definitions of disease codes for example.

Authors: [Juho Kopra](https://github.com/jukop), [Jani Miettinen](https://github.com/janikmiet), [Reijo Sund](https://github.com/rsund)

## Installation

You can install the released version of regstudies from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("regstudies") # maybe some day!! 
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("pohjois-savon-tietoallas/regstudies")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(regstudies)
## basic example code

# ´regstudies´ makes it easy to classify codes such as ICD-codes to groups and also to sum scores based on the groupings.
sample_data <- left_join(sample_cohort,sample_regdata)
sample_data %>% 
  classify_charlson(icd_codes = CODE1) %>%
  sum_score(score_charlson)

# Users can define the code groups themselves using simple regular expression syntax.
my_classes <- head(charlson_classes,5) %>%
  mutate(score_myscore=1:5)

# This classification definition has four different type of codes.
head(charlson_classes)

# Data has three different types of codes.
head(sample_data)

# Different groups types can are automatically handled simultaneously.
sample_data %>% 
  classify_codes(codes=CODE1,diag_tbl = read_classes(my_classes)) %>%
  sum_score(score_myscore)


```


## Citation

To cite package ‘regstudies’ in publications use:

  Juho Kopra, Jani Miettinen and Reijo Sund (2020). regstudies: Tools for
  analyzing register data. R package version 0.5.1.

A BibTeX entry for LaTeX users is

```
  @Manual{,
    title = {regstudies: Tools for analyzing register data},
    author = {Juho Kopra and Jani Miettinen and Reijo Sund},
    year = {2020},
    note = {R package version 0.5.1},
  }

```



For developers:

- [release notes](http://r-pkgs.had.co.nz/release.html)
- [pkgdown](https://pkgdown.r-lib.org/articles/pkgdown.html)
- [pkg dev](http://r-pkgs.had.co.nz/)

