---
title: "Data wrangling for register studies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Data wrangling for register studies}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load necessary libraries for this example

```{r, warning=FALSE, message=FALSE}
library(regstudies) 
library(tidyverse)
library(vroom)
library(lubridate)
```


We need to also create a new help variable `icd` which determines if code is ICD-9, ICD-10 or some other format by year:

```{r}
d <- left_join(sample_cohort,sample_regdata)

```

# Data wrangling for register data sets

### Example data set

The regstudies package also provides useful tools for modifying and filtering the register data sets. Typically in register studies the data consists of events, which are stored in data as time stamps. In our example data set the time stamps are date of hospital visits. Also, in register studies the data are commonly studied with respect to index date, which is a date of interest that can vary by individuals.

Our example data is simulated data that represents a postal questionnaire survey study which has been extended to register data. There are two data sets, the `sample_cohort` is holding the study group, and `sample_regdata` holds the hospital visits for the study group members.

More precisely, the `sample_cohort` lists the individual id study numbers as `id` for whom the register data has been collected. It also list the `postingdate` holding the date of submitting the postal questionnaire.
The `sample_regdata` contains also variable `id` which is necessary to be able to link the data sets. The `sample_regdata` also contains variables holding the diagnosis codes at `CODE1` and times of hospital admission `adm_date` and hospital discharge `disc_date`.

```{r}
head(sample_cohort)
```

```{r}
head(sample_regdata)
```
# Filtering

## Filtering by dates

Filter dataset with either `adm_date` or `disc_date` within two years around `postingdate`

```{r}
filtered_d <- d %>% 
  regstudies::filter_date_range(indexdate=postingdate, 
                          range=years(2),
                          adm_date,
                          disc_date)
head(filtered_d)
```

## Filtering by date intervals

Filter data set with occurrences of hospital visits. If person has been in hospital during the time period two years prior `postingdate` then those events are included.


```{r}
filtered_d <- d %>% 
  filter_date_interval(indexdate = postingdate,
                       time_before=years(2),
                       admission_date=adm_date,
                       discharge_date=disc_date)
head(filtered_d)
```

### Complex filtering

```{r}
elixh_d <- d %>% 
  classify_elixhauser(CODE1)
head(elixh_d)
```

### Left join with filling missing values

In register studies a situation often occurs where the `left_join` creates missing data to to data. This can happen for example when joining `sample_cohort` to `sample_regdata` as `left_join(sample_cohort,sample_regdata)` when there are individuals in cohort who do not hold any records in the register data set. In sample data there are 100 such individuals. Usually this is just a nuisance and we want to fill those values with 0 or false or what ever suits our use case.

For this problem there is function `left_join_replace_na` for which user can give a list of default values what he or she wants to use.

```{r}
d<-left_join(sample_cohort,sample_regdata) %>%
  arrange(personid)
summary(d)
```
This time the missing values are filled based on the defaults entered to `na_replace_list`.

```{r}
na_replace_list <- list(
 "logical"=FALSE,
 "character"="",
 "integer"=0L,
 "double"=0.0,
 "numeric"=0,
 "Date"=lubridate::dmy("01-01-1900")
)
d_filled<-left_join_replace_na(sample_cohort,sample_regdata,na_replace_list)
summary(d_filled)
```

Internally the filling of missing values with default values is executed by another function `replace_na_by_type` which is also available for users.

```{r}
## Give list of default fill values overridding NA
d_filled <- d %>%
 mutate_all(replace_na_by_type,na_replace_list)
summary(d_filled)
```

