---
title: "Example of regstudies workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Example of regstudies workflow}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load necessary libraries for this example

```{r, warning=FALSE, message=FALSE}
library(regstudies) 
library(tidyverse)
library(vroom)
library(lubridate)
```



## Data preparation

For demonstration purpose we are going to create random cohort and registry datasets

```{r}
## Generate random cohort id data
cohort_data <- tibble::tibble(
  personid = seq(1201, 1500),
  gender = sample(c(1,2), 300, replace = T),
  postingdate = rep(as.Date("2000-01-01"), 300)
)

## Generate random icd data
random_icd_codes <- c(
  paste0("F0", seq(1,9)),
  paste0("F", seq(10,48)),
  paste0("F", seq(50,98)),
  paste0("I0", seq(0,2)),
  paste0("I0", seq(05,09)),
  paste0("I", seq(10,16)),
  paste0("I", seq(20,25)),
  paste0("I", seq(30,52)),
  paste0("I", seq(60,99)),
  paste0("E0", seq(0,7)),
  paste0("E0", seq(08,13)),
  paste0("E", seq(15,16)),
  paste0("E", seq(20,35)),
  paste0("E", seq(40,46)),
  paste0("E", seq(70,89))
)
reg_data <- tibble::tibble(
  personid = sample(cohort_data$personid, 10000, replace = TRUE),
  CODE1 = sample(random_icd_codes, 10000, replace = TRUE),
  adm_date = sample(seq(as.Date("1990-01-01"), as.Date("2005-12-31"), by = 1), 10000, replace = T)
  # disc_date = ,
)
reg_data$disc_date <- (reg_data$adm_date + sample(seq(0,180), 10000, replace = T))
```


### Sample dataset 

Dataset `cohort_data` is base dataset including basic descriptive variables of the population (in this case only gener) and variable `posting_date` which is a survey letter posting date:

```{r}
summary(cohort_data)
```

Dataset `reg_data` is a dataset which has events for the selected population with ICD codes and date-variables:

```{r}
summary(reg_data)
```


Joining these two datasets we got dataset ready for analyzes:

```{r}
# Join register data with population cohort data
d <- reg_data %>%
  left_join(cohort_data)

head(d)
```

We can also create a new help variable `icd` which determines if code is ICD-8, ICD-9 or ICD-10 by year:

```{r}
# Determine if codes are ICD-8, ICD-9 or ICD-10
d <- d %>%
  mutate(icd = case_when(
    year(disc_date) < 1987 ~ "icd8",
    year(disc_date) < 1996 & year(disc_date)>=1987 ~ "icd9",
    year(disc_date) >= 1996 ~ "icd10"
    ))
```


## Get first event per id

With tidyverse verbs you can slice first occurence by

```{r}
d %>% 
  group_by(personid) %>% 
  arrange(adm_date) %>%
  dplyr::slice(1)
```

and with regstudies verbs you can do the same with

```{r}
d %>% 
  group_by(personid) %>%
  regstudies::get_first(var = adm_date)

```




## Filtering by date periods

Filter dataset with occurences of hospital visits. If person has been in hospital during the time period two years prior `postingdate` then those events are included.

```{r}
filtered_d <- d %>% 
  filter_date_interval(indexdate = postingdate,
                       time_before=years(2),
                       admission_date=adm_date,
                       discharge_date=disc_date)
head(filtered_d)
```

Filter dataset with either `adm_date` or `disc_date` within two years around `postingdate`

```{r}
filtered_d <- d %>% 
  regstudies::filter_date(indexdate=postingdate, 
                          range=years(2),
                          adm_date,
                          disc_date) 
head(filtered_d)
```


## Classifications

### Code classification by own csv table



```{r}
# Classify by Elixhauser:
my_classes <- read_classes_csv(file = "../data/classification_codes/elixhauser_classes.csv")
classified_d <- filtered_d %>%
  classify_codes(icdcodes = CODE1,
                 diag_tbl = my_classes) 
head(classified_d)
```


### Elixhauser classification and scores

This calculates per row elixhauser classification. Check more info of classification from [classification tables](docs/articles/classification_tables.html)-section.

```{r}
elixscore <- filtered_d %>% 
  classify_elixhauser(icd_codes = CODE1)
head(elixscore)
```

```{r}
### TODO KIRJOITA AUKI
charlson <- filtered_d %>% 
  classify_charlson(icd_codes = CODE1)

## Funktio oletuksena
d %>% 
  classify_codes() -> long # ns. tidy standardi

## ok, kaytaja haluaa leveassa muodossa
## esimerkkiin
d %>% 
  classify_codes() %>% 
  dplyr:: pivot_longer() -> wide
```



Calculate Elixhauser scores by id's

```{r}
# Calculate scores
filtered_d %>% 
  classify_elixhauser(icd_codes = CODE1) %>% 
  sum_score(score_AHRQ, score_van_Walraven)
```





# TODO

## Elixhauser continues

```{r,eval=FALSE}
elixscore %>%
  get_var_types()

elixscore
elixscore <- left_join_replace_na(cohort %>% select(personid), elixscore)
#View(elixscore)

# [X] demoa Hospital Frailty Risk Score -indeksiä!
# [X] selvitä, mitkä funktiot on tarpeettomia ja mitkä tarpeellisia
# TODO: 
# - class, label ja score muotoon class_charlson, label_charlson ja score_charlson
#   -> classify_data_long -funktiossa!
# - korjaa classify_data_long -funktio sellaiseksi että se palauttaa koko datan!
# - tarvitaan left_join, joka tekee muuttujan tyypin mukaan käyttäjän haluaman täytön: esim. c("numeric"=0,"character"="Elixhauser")
#   -> tulee ehkä tarpeettomaksi, jos muutetaan class, label ja score uusille nimille

```



## Charlson classification

```{r, eval=FALSE}
# Charlson score:
#charlson_classes <- classes_to_wide(vroom(file = "data/classification_codes/charlson_classes_wide.csv"))
charlson_classes <- read_classes_csv(file = "data/classification_codes/charlson_classes_wide.csv")
#View(charlson_classes)

# dev:
dev1<-filtereddata %>%
  classify_data_long(icdcodes=CODE1,diag_tbl=charlson_classes)
dev2<-filtereddata %>%
  classify_long(icdcodes=CODE1,diag_tbl=charlson_classes)
dim(filtereddata)
dim(dev1)
dim(dev2)
# TODO: Jotain kummaa tässä on!

charlsonscore <- filtereddata %>%
  classify_data_long(icdcodes=CODE1,diag_tbl=charlson_classes) %>%
  rename(score_charlson=score) %>%
  sum_score(score_charlson)
charlsonscore <- left_join(cohort %>% select(personid), charlsonscore)
#View(charlsonscore)

# Hospital Frailty Risk Score:
hfrs_dat<-vroom(file = "data/classification_codes/hospital_frailty_risk_score_wide.csv")
hfrs_dat<-hfrs_dat %>%
  mutate(label="Frail group") %>%
  select(classification,class,label,score,icd10)
names(hfrs_dat)
write.table(hfrs_dat,file="data/classification_codes/hospital_frailty_risk_score_wide_v2.csv",sep=";",dec=".",row.names = FALSE)
View(hfrs_dat)
hfrs_classes <- read_classes_csv(file = "data/classification_codes/hospital_frailty_risk_score_wide_v2.csv")
#View(charlson_classes)
hfrsscore <- filtereddata %>%
  classify_data_long(icdcodes=CODE1,diag_tbl=hfrs_classes) %>%
  rename(score_charlson=score) %>%
  sum_score(score_charlson)

mode <- function(x) { # by Ken Williams at stackoverflow
  ux <- unique(x)
  ux <- ux[!is.na(ux)]
  ux[which.max(tabulate(match(x, ux)))]
}
mode(hfrsscore$classification)


hfrsscore <- left_join(cohort %>% select(personid), hfrsscore) %>%
  mutate(classification=mode(classification))
#  mutate(classification=ifelse(is.na(classification),mode(classification),classification))
View(hfrsscore)

dim(filtereddata %>% select(personid) %>% distinct())
dim(cohort)
dim(elixscore)
dim(charlsonscore)

bothscores<-left_join(elixscore %>% select(-classification),
                      charlsonscore %>% select(-classification),by="personid")
fill_na<-function(x,fill=0) ifelse(is.na(x),fill,x)
cohortscores<-left_join(cohort %>% select(personid),bothscores) %>%
  mutate_all(fill_na)
cohortscores %>%
  head(10)
cohortscores %>%
  filter(`sum(score_AHRQ)`+`sum(score_van_Walraven)`+`sum(score_charlson)`>0) %>%
  head(10)
plot(cohortscores)
cor(cohortscores)
```

